\name: Deterministic and Coherent Implicits

### 2017-02-27 src/Translation.tex
Section 4

Moreover, we can express three key properties of Figure~\ref{fig:resolution2}'s
definition of resolution in terms of the generated evidence.
\begin{lemma}[Determinacy]
The generated evidence of resolution is uniquely determined.
\[\forall \tenv, \rulet, E_1, E_2: \quad\quad \tenv \ivturns \rulet \leadsto E_1 ~\wedge~ \tenv \ivturns \rulet \leadsto E_2 \quad\Rightarrow\quad E_1 = E_2 \]
\end{lemma}
\begin{lemma}[Soundness]
Figure~\ref{fig:resolution2}'s definition of resolution is sound (but
incomplete) with respect to Figure~\ref{fig:resolution1}'s definition.
\[\forall \tenv, \rulet, E: \quad\quad \tenv \ivturns \rulet \leadsto E \quad\Rightarrow\quad \tenv \vturns \rulet \leadsto E \]
\end{lemma}
\begin{lemma}[Coherence]
Resolution is stable under substitution.
\[\forall \tenv,\alpha,\tenv',\sigma,\rulet, E: \quad\quad 
\tenv,\alpha,\tenv' \ivturns \rulet \leadsto E \quad\wedge\quad \tenv \vdash \sigma
\quad\Rightarrow\quad 
\tenv,\tenv'[\sigma/\alpha] \ivturns \rulet[\sigma/\alpha] \leadsto E[|\sigma|/\alpha] \]
\end{lemma}

@@

The evidence generation in Figure~\ref{fig:algorithm} is largely similar to
that in the deterministic specification of resolution in
Figure~\ref{fig:resolution2}.

@@

With the evidence we can state the correctness of the algorithm.
\begin{theorem}[Partial Correctness]
Let $\tenv$ be a type environment, $\rulet$ be a type and E be a System F expression.
Assume that $\epsilon \vdash_{\mathit{unamb}} \rulet$ and also $\forall \rulet_i \in \tenv: \epsilon \vdash_{\mathit{unamb}} \rulet_i$.
Then $\tenv \ivturns \rulet \leadsto E$ if and only if $\tenv \vdash_{\mathit{alg}} \rulet \leadsto E$,
provided that the algorithm terminates.
\end{theorem}

>>>>

Moreover, we can express three key properties of Figure~\ref{fig:resolution2}'s
definition of resolution in terms of the generated evidence.
Firstly, the deterministic version of resolution is a sound variation on the original ambiguous resolution.
\begin{lemma}[Soundness]
Figure~\ref{fig:resolution2}'s definition of resolution is sound (but
incomplete) with respect to Figure~\ref{fig:resolution1}'s definition.
\[\forall \tenv, \rulet, E: \quad\quad \tenv \ivturns \rulet \leadsto E \quad\Rightarrow\quad \tenv \vturns \rulet \leadsto E \]
\end{lemma}
Secondly, the deterministic resolution guarantees a strong form of coherence:
\begin{lemma}[Determinacy]
The generated evidence of resolution is uniquely determined.
\[\forall \tenv, \rulet, E_1, E_2: \quad\quad \tenv \ivturns \rulet \leadsto E_1 ~\wedge~ \tenv \ivturns \rulet \leadsto E_2 \quad\Rightarrow\quad E_1 = E_2 \]
\end{lemma}
Thirdly, on top of the immediate coherence of deterministic resolution, 
an additional coherence property holds.  
\begin{lemma}[Stability]
Resolution is stable under substitution.
\[\forall \tenv,\alpha,\tenv',\sigma,\rulet, E: \quad\quad 
\tenv,\alpha,\tenv' \ivturns \rulet \leadsto E \quad\wedge\quad \tenv \vdash \sigma
\quad\Rightarrow\quad 
\tenv,\tenv'[\sigma/\alpha] \ivturns \rulet[\sigma/\alpha] \leadsto E[|\sigma|/\alpha] \]
\end{lemma}

@@

The evidence generation in Figure~\ref{fig:algorithm} is largely similar to
that in the deterministic specification of resolution in
Figure~\ref{fig:resolution2}.
With the evidence we can state the correctness of the algorithm.

@@

\begin{theorem}[Partial Correctness]
Let $\tenv$ be a type environment, $\rulet$ be a type and E be a System F expression.
Assume that $\epsilon \vdash_{\mathit{unamb}} \rulet$ and also $\forall \rulet_i \in \tenv: \epsilon \vdash_{\mathit{unamb}} \rulet_i$.
Then $\tenv \ivturns \rulet \leadsto E$ if and only if $\tenv \vdash_{\mathit{alg}} \rulet \leadsto E$,
provided that the algorithm terminates.
\end{theorem}
